Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-23 cav2017/cav2017.tex
move extensions to appendix
uurghhg fixed point arithmetic section is a mess

The 
%systems and 
models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e. to represent the plant dynamics $A_d$, $B_d$ and its states. 
The second is the numeric error introduced by the digital controller, 
which 
%uses an analogue-to-digital converter to read the states as a fixed-point number and performs operations in this domain. 
performs operations in the domain of fixed-point numbers. 
In this section we formally outline the errors introduced by fixed-point representation of numbers. 
In further sections we will differentiate between these two errors, 
by referring to them as the \emph{plant error} and the \emph{digital controller error} respectively.

@@

Thus far we have presented a verification procedure for a digitalised plant in a continuous space (\emph{ie} the
range of the quantisers is still $\mathbb{R}$. The next step we must realise is that the controller 
exists in a digital program, hence its domain is $\mathbb{R}\langle I,F\rangle$. We have $\mat{A},\mat{B},\vec{x} \in \mathbb{R}^*$, $\mat{K}_d,\vec{u},\hat{\vec{x}}=\mathcal{F}\langle I,F \rangle(\vec{x}) \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix/vector.
Unfortunately this means that although the closed loop dynamics are in the reals, they suffer from
the fixed word restrictions in their composing elements.
Apart from the quantisation errors $q_1,q_2$ there are rounding errors in the calculation of $u_k$. 
These will introduce a new source of quantisation noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=(|\mat{K}|_1+1)c_m$

@@

\subsubsection{Effect on safety specification and stability}
The states values must be computed considering the digital controller error, i.e., the error introduced by the ADC and the digital controller. 
%, and the digital reference signal. 
%In this paper, we consider systems with a reference signal equal to zero. 
%the FWL effects in all operations related to the control action~\eqref{eq:controlaction} 
%and disregarding it in the state equation~\eqref{eq:plant}.
Thus, the states are computed using the following equations. We use $\mathcal{F}_{\langle I_c,F_c \rangle}$ \todo{explain and clarify notation.}
to to denote numbers represented with the precision of the digital controller. All other numbers
are represented with the precision of our plant model:
%
\begin{align*}
u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})).  
%\\ x_{k+1}&=A_d x_{k} + B_du_{k}
\end{align*}

@@

LTI models in the form: 
%
\begin{align}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\ \ :\ \ \vec{x} \in \mathbb{R}^{n}, \vec{u} \in \mathbb{R}^m, \mat{A} \in \mathbb{R}^{n \times n},\mat{B} \in \mathbb{R}^{n \times m}\\
\vec{y}(t)&=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t)\ \ :\ \ \vec{y} \in \mathbb{R}^{o}, \mat{C} \in \mathbb{R}^{o \times n}, \mat{D}  \in \mathbb{R}^{o \times m}\nonumber
\end{align}
\noindent where $\dot{\vec{x}}(t)$ describes the state evolution equation; 
$\vec{y}(t)$ represents an instantaneous output equation; and $\mat{A}$, $\mat{B}$, $\mat{C}$, and $\mat{D}$ are matrices that fully specify 
a continuous plant. Eq.~\eqref{eq:ode} is soundly discretized 
(as shown in the appendix) into
%
\begin{align}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k .\nonumber
\end{align}
%

@@

\end{document}

>>>>

The 
%systems and 
models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e. to represent the plant dynamics $A_d$, $B_d$ and its states. 
The second is the quantization error introduced by the digital controller, 
which 
performs operations in the domain of fixed-point numbers. 
In this section we formally outline the errors introduced by fixed-point representation of numbers. 
In further sections we will differentiate between these two errors, 
by referring to them as the \emph{plant error} and the \emph{digital controller error} respectively.

@@

Thus far we have presented a verification procedure for a digitalised plant in a continuous space (\emph{i.e.} the
range of the quantisers is still $\mathbb{R}$. The next step we must realise is that the controller 
exists in a digital program, hence its domain is $\mathbb{R}\langle I,F\rangle$. We have $\mat{A},\mat{B},\vec{x} \in \mathbb{R}^*$, $\mat{K}_d,\vec{u},\hat{\vec{x}}=\mathcal{F}\langle I,F \rangle(\vec{x}) \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix/vector.
Unfortunately this means that although the closed loop dynamics are in the reals, they suffer from
the fixed word restrictions in their composing elements.
Apart from the quantisation errors $q_1,q_2$ there are rounding errors in the calculation of $u_k$. 
These will introduce a new source of quantisation noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=(|\mat{K}|_1+1)c_m$

@@

\subsubsection{Effect on safety specification and stability}
First consider the effect of the quantization error on safety; the state values must be computed considering the error introduced by the digital controller caused by performing the multiplication $K*x$ at low digital precision.
%, and the digital reference signal. 
%In this paper, we consider systems with a reference signal equal to zero. 
%the FWL effects in all operations related to the control action~\eqref{eq:controlaction} 
%and disregarding it in the state equation~\eqref{eq:plant}.
Thus, the states are computed using the following equations. We use $\mathcal{F}_{\langle I_c,F_c \rangle}$ 
to to denote numbers represented with the precision of the digital controller, where $I_c$ is the number of bits used to represent the integer value in the controller and $F_c$ is the number of bits used to represent the fractional part of numbers in the controller. 
%
\begin{align*}
u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})).  
%\\ x_{k+1}&=A_d x_{k} + B_du_{k}
\end{align*}

@@

%-------------------------------
\section{Conclusions}
%-------------------------------

@@

LTI models in the form: 
%
\begin{align}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\ \ :\ \ \vec{x} \in \mathbb{R}^{n}, \vec{u} \in \mathbb{R}^m, \mat{A} \in \mathbb{R}^{n \times n},\mat{B} \in \mathbb{R}^{n \times m}\\
\vec{y}(t)&=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t)\ \ :\ \ \vec{y} \in \mathbb{R}^{o}, \mat{C} \in \mathbb{R}^{o \times n}, \mat{D}  \in \mathbb{R}^{o \times m}\nonumber
\end{align}
\noindent where $\dot{\vec{x}}(t)$ describes the state evolution equation; 
$\vec{y}(t)$ represents an instantaneous output equation; and $\mat{A}$, $\mat{B}$, $\mat{C}$, and $\mat{D}$ are matrices that fully specify 
a continuous plant. Eq.~\eqref{eq:ode} is soundly discretized 
(as shown in the appendix) into
%
\begin{align}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k .\nonumber
\end{align}
%
\end{document}
